Handling missing data in smartphone location logs

### 2017-11-17 thesisReport.Rmd
added plots and table

Borrowing from @barnett_inferring_2016 let a persons' true Cartesian location be $G(t) = [G_x(t) G_y(t)]$ and $D \in \mathbb{R}^2$ be the recorded data containing lattitude and longitude. In addition, let $a$ denote the estimated accuracy of the recorded data.  accuracy. $G(t)$, $D$ and $a$ are indexed by the set $t_1 < ... < t_{n+1}$. For simplicity, let each entry in the discrete index set represent a 5 minute window. The measure of accuracy $a_t$ is given in meters such that it represents the radius of a 67% confidence circle. If $D_t = \emptyset$ it is considered *missing* and it is not missing otherwise.

@@

Given that Android phones collect location information from WiFi access points, cellphone triangulation, and GPS measurements due to computational and battery constraints [@lamarca_place_2005; @chen_practical_2006], the accuracy is substantially lower than in professional grade GPS trackers. Based on this data set, less than 40% of measures $D$ are paired with $a \leq 20$. @palmius_detecting_2017 note that in their custom logs inaccurate location values are interspersed between more accurate location values at higher sample rates per hour. I observe similar patterns in secondary logs.

@@

@@

```{r datadetailsTable, echo = F}
datadetails <- readRDS("tempdata/datadetails.rds")
kable(datadetails,digits = 2,col.names = c("Log duration",
                                           "Observations",
                                           "Missing days",
                                           "Mean Accuracy",
                                           "SD Accuracy"))
```

>>>>

Borrowing from @barnett_inferring_2016 let a persons' true Cartesian location on an error minimising projection be $G(t) = [G_x(t) G_y(t)]$ and $D \in \mathbb{R}^2$ be the recorded data containing lattitude and longitude. In addition, let $a$ denote the estimated accuracy of the recorded data.  accuracy. $G(t)$, $D$ and $a$ are indexed by the set $t_1 < ... < t_{n+1}$. For simplicity, let each entry in the discrete index set represent a 5 minute window. The measure of accuracy $a_t$ is given in meters such that it represents the radius of a 67% confidence circle. If $D_t = \emptyset$ it is considered *missing* and it is not missing otherwise.

@@

Given that Android phones collect location information from WiFi access points, cellphone triangulation, and GPS measurements due to computational and battery constraints [@lamarca_place_2005; @chen_practical_2006], the accuracy is substantially lower than in professional grade GPS trackers. Based on this data set the distribution of $a$ is highly right skewed, with a median of 28, $\mu = 127$ and the maximum value at 26 km. @palmius_detecting_2017 note that in their custom logs inaccurate location values are interspersed between more accurate location values at higher sample rates per hour. I observe similar patterns in secondary logs.

```{r, accuracyPlot2, echo=FALSE, fig.cap="The accuracy of raw measures with time on the x-axis and accuracy on the y-axis. The colour scale shows the distance between the measurement and the previous point. To ease interpretability all points with a distance between points of less than 5 meters have been filtered out.",out.width = '100%'}
include_graphics("img/accuracyLocShift.png")

@@

In our example data set, over 54% of the data is missing for the entire duration of the log. However,this is misleading as there are several long periods with no measurements whatsoever. The typical strucutre of missingness of a day with measurements is shown below.

@@

```{r datadetailsTable, echo = F, fig.cap="caption here?"}
datadetails <- readRDS("tempdata/datadescriptives.rds")
datadetails <- datadetails[,-7]
kable(datadetails,digits = 2,col.names = c("Log duration",
                                           "Days of log",
                                           "Observations",
                                           "Missing days",
                                           "Missing data",
                                           "Mean Accuracy"))
```
